On the Synchronization Power of the OpenFlow Data Plane

### 2015-05-22 inband-sync.tex
partial Sec 5

\newcommand{\error}{\texttt{error}}
\newcommand{\xid}{\texttt{xid}}
\newcommand{\ecode}{\texttt{error-code}}
\newcommand{\exec}{\textbf{execute}}
\newcommand{\execatomic}{\textbf{execute-atomic}}
\newcommand{\ack}{\textit{ack}}

@@

\section{Use Cases}\label{sec:apps}

@@

\subsection{Compare-and-Set}
The pathological scenario described in the example in
Figure~\ref{fig:CAS-example} can be easily resolved if we put
policy update in a transactional $\execatomic}$ construct together
 with a \emph{compare-and-set} (CAS) operation on the policy identifier.
A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$ 
checks if the content of $\addr$ is $\textit{old}$ and if so, 
replaces it with $\textit{new}$ and returns $\true$ (the CAS
\emh{succeeds}), otherwise it returns $\true$ (the CAS
\emh{fails} (which causes an abort of
the invoking transaction). 
It is straightforward to implement the CAS operation by executing
$\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
new)$ withing the same bundle.

@@

    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}
%]]

CAS can be used to implement a wide range powerful abstractions
shared by the controllers, such as queues, sets, hash tables,
etc.~\cite{Her91} in the \emph{wait-free} manner, i.e., tolerating
asynchrony and failure of any number of controllers.      
In our transactional setting, the controllers
can use CAS to replace the default value of such a register
with their consensus input values (e.g., the identifiers of the
policies that should be applied first). If the corresponding
singleton transaction commits, the controller considers itself
the \emph{winner} and decides on its input, if
it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
winner to decide.

By using multiple CAS invocations, the controllers can implement any
generic replicated state machine~\cite{Her91,Lam98}, which will be instrumental in the
next section.

Moreover, given the CAS, we can also implement more complex operations,
for example the replicated state machine postulated in STN~\cite{stn}:
in order to choose tags in a consistent manner, CAS needs to be enhanced
with a notion of tags.

\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}

\petr{"more complex operations": more complex than what?}

\petr{the discussion of CAS in STN is confusing, given that we never described STN.}

\subsection{Policy Updates}

>>>>

\newcommand{\error}{\texttt{error}}
\newcommand{\xid}{\texttt{xid}}
\newcommand{\ecode}{\texttt{error-code}}
\newcommand{\exec}{\textbf{execute}}
\newcommand{\execatomic}{\textbf{execute-atomic}}
\newcommand{\ack}{\textit{ack}}
\newcommand{\true}{\textit{true}}
\newcommand{\false}{\textit{false}}

@@

\section{Applications}\label{sec:apps}

@@

\paragraph{Compare-and-Set}
The pathological scenario described in the example in
Figure~\ref{fig:CAS-example} can be easily resolved if we put
policy update in our transactional $\execatomic$ construct together
 with a \emph{compare-and-set} (CAS) operation on the policy identifier.
A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$ 
checks if the content of $\addr$ is $\textit{old}$ and if so, 
replaces it with $\textit{new}$ and returns $\true$ (the CAS
\emph{succeeds}), otherwise it returns $\true$ (the CAS
\emph{fails} (which causes an abort of
the invoking transaction). 
It is straightforward to implement the CAS operation by executing
$\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
new)$ withing the same bundle.

@@

    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}
}
%]]

A single CAS invocation can be used to solve the fundamental
problem of fault-tolerant \emph{consensus}~\cite{FLP85} in which the controllers
need to agree on one of their private inputs, despite failures of an
arbitrary number of controllers.    
Indeed, each controller simply performs, within a transaction, a CAS operation to replace the default value
in a memory location at the meta-configuration of a data-plane switch.
If the transaction commits, the controller considers itself
the \emph{winner} and decides on its input, 
if it aborts (the controller fails in the CAS), it remains to read the
memory to get the input of the winner and decide on it.

Using multiple CAS invocations, the controllers can implement any
generic replicated state machine~\cite{Her91,Lam98} on the control plane, 
in the \emph{wait-free} manner, i.e., tolerating
asynchrony and failure of any number of controllers.
%which will be instrumental in the
%next section.

%Moreover, given the CAS, we can also implement more complex operations,
%for example the replicated state machine postulated in STN~\cite{stn}:
%in order to choose tags in a consistent manner, CAS needs to be enhanced%
%with a notion of tags.

%\petr{when talking about replicated state machines, we should make it clear that we that we talk about replication among controllers.}

%\petr{"more complex operations": more complex than what?}

%\petr{the discussion of CAS in STN is confusing, given that we never described STN.}

\paragraph{Policy Updates}
